---
title: "Example tidyphreeqc usage"
author: "Dewey Dunnington"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
# library(tidyverse) causes an CMD check note,
# and depending on tidyverse is a bad idea according HW.
library(ggplot2)
library(tidyr)
library(dplyr)
library(stringr)

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Example

Running PHREEQC is accomplished using the `phr_run()` function, which calls the program and generates the output. The function accepts character vectors of input, which can be generated using intput helper functions such as `phr_solution()`, `phr_selected_output()`, `phr_equilibrium_phases()`, and `phr_reaction_temperature()` (or roll your own input using `phr_input_section()`).

```{r}
library(tidyphreeqc)
phr_run(
  phr_solution(pH = 7, temp = 25)
)
```

To get the results as a data frame, we need to supply a `phr_selected_output()` to the input file.

```{r}
phr_run(
  phr_solution(pH = 7, temp = 25),
  phr_selected_output(pH = TRUE, temp = TRUE, activities = c("OH-", "H+", "O2"))
)
```

To find the distribution of a few solutions, you can generate a list of solutions using `phr_solution_list()`.

```{r}
phr_run(
  phr_solution_list(pH = 5:8, temp = 12:25),
  phr_selected_output(pH = TRUE, temp = TRUE, activities = c("OH-", "H+", "O2"))
)
```

## Databases

Some elements (for example, mercury) aren't included in the base database. There are a number of databases included in the PHREEQC package, that you can choose by specifying the `db` argument of `phr_run()`. One that includes mercury is the "minteq" database.

```{r}
phr_run(
  phr_solution(pH = 7, temp = 25, Hg = 0.1),
  phr_selected_output(
    activities = c("Hg", "Hg2+2", "Hg(OH)2", "Hg(OH)2", "HgOH+", "Hg(OH)3-")
  ),
  db = "minteq"
)
```

## Pourbaix diagrams

```{r}
result <- phr_run(
  phr_solution_list(pH = seq(0, 14, 0.5), pe = seq(-14, 22, 0.5), Hg = 0.1),
  phr_selected_output(
    activities = c("Hg", "Hg2+2", "Hg(OH)2", "Hg(OH)2", "HgOH+", "Hg(OH)3-")
  ),
  db = "minteq"
)

result
```

```{r pourbaix}
# library(tidyverse)
result_long <- result %>%
  as_tibble() %>%
  gather(key = "species", value = "log_activity", starts_with("la_")) %>%
  mutate(species = str_remove(species, "^la_"))

result_long %>%
  ggplot(aes(x = pH, y = pe, fill = log_activity)) +
  geom_raster() +
  stat_contour(aes(z = log_activity)) +
  facet_wrap(~species)

result_long %>%
  filter(species != "Hg") %>%
  group_by(pH, pe) %>%
  summarise(
    dominant_species = species[which.max(log_activity)],
    dominant_log_act = max(log_activity),
    dominance = max(log_activity) - max(setdiff(log_activity, max(log_activity)))
  ) %>%
  ggplot(aes(pH, pe, col = dominant_species, alpha = dominance)) +
  geom_point()
```

## Getting raw info

```{r}
phr_solution(Hg = 0.1) %>% 
  phr_run(db = "minteq", dump = TRUE) %>% 
  phr_print_output()
```

```{r}
phr_solution(Hg = 0.1) %>% 
  phr_run(db = "minteq", dump = TRUE) %>% 
  phr_print_dump()
```
